---
title: "seminrExtras 1.0.0 — Feature Tour"
author: "Nicholas P. Danks"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

`seminrExtras` 1.0.0 is now on CRAN. It extends the `seminr` PLS-SEM workflow with eight new assessment methods:

| Method    | Function(s)                                           | What it tells you                                     |
|-----------|-------------------------------------------------------|-------------------------------------------------------|
| COA       | `assess_coa()`                                        | Why / for whom your model fails to generalise         |
| NCA       | `assess_nca()`                                        | Which predictors are **necessary** (not just sufficient) |
| NCA-ESSE  | `assess_nca_esse()`                                   | How robust NCA conclusions are to extreme responses    |
| cIPMA     | `assess_cipma()` / `assess_ipma()`                    | Managerial priority map, optionally with necessity    |
| FIMIX-PLS | `assess_fimix()` / `assess_fimix_compare()`           | EM-based latent-class segmentation                    |
| PLS-POS   | `assess_pos()` / `assess_pos_compare()` / `pos_segments()` | Prediction-oriented segmentation (max ΣR²)        |
| CTA-PLS   | `assess_cta()`                                        | Is this measurement model actually reflective?        |
| PCM       | `assess_pcm()`                                        | Does the mediator improve out-of-sample prediction?    |

Plus hardened `assess_cvpat()`, `assess_cvpat_compare()`, and `congruence_test()`.

```{r libs}
# install.packages("seminrExtras")
library(seminr)
library(seminrExtras)
packageVersion("seminrExtras")
```

Two bundled datasets are used throughout:

- `mobi` — MOBI customer satisfaction (from `seminr`)
- `corp_rep_data` — Corporate reputation (from `seminr`)

---

# 1. Base models

## 1a. MOBI model (reflective, for NCA / cIPMA / CTA / PCM / congruence)

```{r mobi-model}
mobi_mm <- constructs(
  composite("Image",        multi_items("IMAG", 1:5)),
  composite("Expectation",  multi_items("CUEX", 1:3)),
  composite("Value",        multi_items("PERV", 1:2)),
  composite("Satisfaction", multi_items("CUSA", 1:3)),
  composite("Loyalty",      multi_items("CUSL", 1:3))
)

mobi_sm <- relationships(
  paths(from = "Image",        to = c("Expectation", "Satisfaction", "Loyalty")),
  paths(from = "Expectation",  to = c("Value", "Satisfaction")),
  paths(from = "Value",        to = "Satisfaction"),
  paths(from = "Satisfaction", to = "Loyalty")
)

mobi_pls <- estimate_pls(
  data              = mobi,
  measurement_model = mobi_mm,
  structural_model  = mobi_sm
)
```

## 1b. Corporate reputation model (Mode B formative, for FIMIX / POS / CVPAT / COA)

```{r corp-model}
corp_rep_mm <- constructs(
  composite("QUAL", multi_items("qual_", 1:8), weights = mode_B),
  composite("PERF", multi_items("perf_", 1:5), weights = mode_B),
  composite("CSOR", multi_items("csor_", 1:5), weights = mode_B),
  composite("ATTR", multi_items("attr_", 1:3), weights = mode_B),
  composite("COMP", multi_items("comp_", 1:3)),
  composite("LIKE", multi_items("like_", 1:3)),
  composite("CUSA", single_item("cusa")),
  composite("CUSL", multi_items("cusl_", 1:3))
)

corp_rep_sm <- relationships(
  paths(from = c("QUAL", "PERF", "CSOR", "ATTR"), to = c("COMP", "LIKE")),
  paths(from = c("COMP", "LIKE"),                 to = c("CUSA", "CUSL")),
  paths(from = "CUSA",                            to = "CUSL")
)

corp_pls <- estimate_pls(
  data              = corp_rep_data,
  measurement_model = corp_rep_mm,
  structural_model  = corp_rep_sm,
  missing           = mean_replacement,
  missing_value     = "-99"
)
```

## 1c. Simple COA model (reduced corporate reputation)

```{r coa-model}
corp_mm_simple <- constructs(
  composite("COMP", multi_items("comp_", 1:3)),
  composite("LIKE", multi_items("like_", 1:3)),
  composite("CUSA", single_item("cusa")),
  composite("CUSL", multi_items("cusl_", 1:3))
)

corp_sm_simple <- relationships(
  paths(from = c("COMP", "LIKE"), to = "CUSA"),
  paths(from = "CUSA",            to = "CUSL")
)

corp_model <- estimate_pls(
  data              = corp_rep_data,
  measurement_model = corp_mm_simple,
  structural_model  = corp_sm_simple,
  missing           = mean_replacement,
  missing_value     = "-99"
)
```

---

# 2. Composite Overfit Analysis (COA)

Diagnoses *why* and *for whom* the model fails to generalise. Returns predictive deviance, a decision tree that groups deviant cases, and unstable-path analysis.

```{r coa}
coa_result <- assess_coa(
  corp_model,
  focal_construct = "CUSL",
  noFolds = 10, reps = 1, cores = 1,
  seed = 123
)

print(coa_result)
summary(coa_result)
```

### Plots

```{r coa-plots}
plot(coa_result, type = "pd")      # predictive deviance distribution
plot(coa_result, type = "tree")    # decision tree
plot(coa_result, type = "groups")  # deviant group highlights
```

### Step-by-step COA

```{r coa-steps}
pd     <- predictive_deviance(corp_model,
                              focal_construct = "CUSL",
                              noFolds = 10, reps = 1, cores = 1,
                              seed = 123)
tree   <- deviance_tree(pd, deviance_bounds = c(0.025, 0.975))
instab <- unstable_params(corp_model, tree$deviant_groups, params = "path_coef")
print(instab)

# Inspect the rules and competing splits for one deviant group
first_group <- names(tree$deviant_groups)[1]
group_rules(first_group, coa_result)             # decision rule for group "A"
competes(tree$group_roots[[first_group]], tree)  # competing splits at that node
```

---

# 3. Necessary Condition Analysis (NCA)

Tests **necessity**, not sufficiency. No external `NCA` package required — CE-FDH and CR-FDH ceiling lines are built in.

```{r nca}
nca_sat <- assess_nca(
  mobi_pls,
  target   = "Satisfaction",
  test.rep = 1000,
  seed     = 123
)

print(nca_sat)
summary(nca_sat)

plot(nca_sat, type = "effects")   # effect size bar plot
plot(nca_sat, type = "scatter")   # ceiling-line scatter plots
```

---

# 4. NCA-ESSE (Effect Size Sensitivity Extension)

Varies the ECDF threshold, removes upper-left observations, and compares the empirical effect size drop against a uniform benchmark (Becker et al., 2026).

```{r nca-esse}
esse_sat <- assess_nca_esse(
  mobi_pls,
  target     = "Satisfaction",
  thresholds = seq(0, 0.05, by = 0.005),
  seed       = 123
)

print(esse_sat)
summary(esse_sat)

plot(esse_sat, type = "sensitivity")   # Fig. 4, Becker et al. (2026)
plot(esse_sat, type = "difference")    # Fig. 6, Becker et al. (2026)
```

---

# 5. cIPMA (Combined Importance–Performance Map Analysis)

`assess_ipma()` is a convenience wrapper for IPMA alone; `assess_cipma()` overlays NCA necessity onto the IPMA map.

## IPMA only

```{r ipma}
ipma_result <- assess_ipma(
  mobi_pls,
  target    = "Loyalty",
  scale_min = 1,
  scale_max = 10
)

print(ipma_result)
summary(ipma_result)
plot(ipma_result, type = "ipma")
```

## cIPMA (IPMA + necessity)

```{r cipma}
cipma_result <- assess_cipma(
  mobi_pls,
  target       = "Loyalty",
  scale_min    = 1,
  scale_max    = 10,
  nca_test.rep = 1000,
  seed         = 123
)

print(cipma_result)
summary(cipma_result)
plot(cipma_result, type = "cipma")                       # IPMA + NCA overlay
plot(cipma_result, importance_metric = "standardized")   # standardized total effects
```

---

# 6. FIMIX-PLS (EM-based latent class segmentation)

Uncovers unobserved heterogeneity. Multi-start EM with information-criteria comparison across K.

## Single K

```{r fimix-single}
fimix_k2 <- assess_fimix(corp_pls, K = 2, nstart = 10, seed = 123)

print(fimix_k2)
summary(fimix_k2)
plot(fimix_k2)
```

## Compare across K

```{r fimix-compare}
fimix_compare <- assess_fimix_compare(
  corp_pls,
  K_range = 2:4,
  nstart  = 10,
  seed    = 123
)

print(fimix_compare)
plot(fimix_compare)   # IC across K
```

---

# 7. PLS-POS (Prediction-Oriented Segmentation)

Maximises Σ R² across segments (Becker et al., 2013). No distributional assumptions, detects heterogeneity in both structural and formative measurement models.

## Single K

```{r pos-single}
pos_k2 <- assess_pos(corp_pls, K = 2, nstart = 10, max_iter = 100, seed = 123)

print(pos_k2)
summary(pos_k2)

plot(pos_k2, type = "segments")
plot(pos_k2, type = "rsquared")
plot(pos_k2, type = "paths")
```

## Extract per-segment models

```{r pos-segments}
seg_models <- pos_segments(pos_k2)
summary(seg_models[[1]])
```

## Compare across K

```{r pos-compare}
pos_compare <- assess_pos_compare(
  corp_pls,
  K_range  = 2:4,
  nstart   = 10,
  max_iter = 100,
  seed     = 123
)

print(pos_compare)
plot(pos_compare)
```

---

# 8. CTA-PLS (Confirmatory Tetrad Analysis)

Tests whether a construct's measurement model is consistent with a reflective specification. With `borrow = TRUE` (default), constructs with 2–3 indicators borrow from structurally connected constructs so they can still be tested.

```{r cta}
cta_result <- assess_cta(mobi_pls, nboot = 5000, seed = 123)

print(cta_result)
summary(cta_result)
plot(cta_result)
```

### Without borrowing (only constructs with ≥ 4 indicators)

```{r cta-no-borrow}
cta_no_borrow <- assess_cta(mobi_pls, nboot = 5000, borrow = FALSE, seed = 123)
print(cta_no_borrow)
```

---

# 9. PCM (Predictive Contribution of the Mediator)

Does the mediator actually improve out-of-sample prediction? Compares Direct-Antecedent (DA) and Earliest-Antecedent (EA) predictions on the isolated mediation sub-model (Danks, 2021). Mediation paths are auto-detected.

## Single-mediator model

```{r pcm-single}
# Use a minimal mediation model: Image -> Satisfaction -> Loyalty (+ direct)
pcm_mm <- constructs(
  composite("Image",        multi_items("IMAG", 1:5)),
  composite("Satisfaction", multi_items("CUSA", 1:3)),
  composite("Loyalty",      multi_items("CUSL", 1:3))
)
pcm_sm <- relationships(
  paths(from = "Image",        to = "Satisfaction"),
  paths(from = "Satisfaction", to = "Loyalty"),
  paths(from = "Image",        to = "Loyalty")
)
pcm_pls <- estimate_pls(mobi, pcm_mm, pcm_sm)

pcm_result <- assess_pcm(
  pcm_pls,
  target  = "Loyalty",
  noFolds = 10,
  reps    = 10
)

pcm_result
summary(pcm_result)
plot(pcm_result)
```

## Multi-mediator model (auto-detects all paths)

```{r pcm-multi}
pcm_full <- assess_pcm(
  mobi_pls,
  target  = "Loyalty",
  noFolds = 10,
  reps    = 10
)

pcm_full
summary(pcm_full)
plot(pcm_full)
```

---

# 10. Hardened CVPAT and congruence testing

## CVPAT (single model + model comparison)

```{r cvpat}
# Alternative structural model: drop CUSL from COMP/LIKE block
alt_sm <- relationships(
  paths(from = c("QUAL", "PERF", "CSOR", "ATTR"), to = c("COMP", "LIKE")),
  paths(from = c("COMP", "LIKE"),                 to = "CUSA"),
  paths(from = "CUSA",                            to = "CUSL")
)

corp_pls_alt <- estimate_pls(
  data              = corp_rep_data,
  measurement_model = corp_rep_mm,
  structural_model  = alt_sm,
  missing           = mean_replacement,
  missing_value     = "-99"
)

# Assess the focal model
cvpat_focal <- assess_cvpat(
  corp_pls,
  testtype  = "two.sided",
  nboot     = 2000,
  seed      = 123,
  technique = predict_DA,
  noFolds   = 10,
  reps      = 10
)
print(cvpat_focal$CVPAT_compare_LM, digits = 3)
print(cvpat_focal$CVPAT_compare_IA, digits = 3)

# Compare focal vs. alternative
cvpat_cmp <- assess_cvpat_compare(
  established_model = corp_pls,
  alternative_model = corp_pls_alt,
  testtype  = "two.sided",
  nboot     = 2000,
  technique = predict_DA,
  seed      = 123,
  noFolds   = 10,
  reps      = 10,
  cores     = NULL
)
print(cvpat_cmp, digits = 3)
```

## Congruence test

Tests whether the congruence coefficients between original PLS composite weights and bootstrapped weights are significantly close to 1.

```{r congruence}
cong_result <- congruence_test(
  mobi_pls,
  nboot     = 2000,
  seed      = 123,
  alpha     = 0.05,
  threshold = 1
)

print(cong_result)
summary(cong_result)
```

---

# Session info

```{r sessioninfo}
sessionInfo()
```
